---
title: "mentalhealthupdated"
output: html_document
date: "2022-12-08"
runtime: shiny
---

```{r}
packages_list <- c("tidyverse", "dplyr","janitor","ggplot2","arm","tidyr","GGally","broom","mgcv","metR","MASS","NHANES","plotly","ggthemes","metR")
to_be_installed <- packages_list[!(packages_list %in% installed.packages()[,"Package"])]
if (length(to_be_installed) > 0) {
  install.packages(to_be_installed)
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(janitor)
library(ggplot2)
library(arm)
library(tidyr)
library(GGally)
library(broom)
library(mgcv)
library(metR)
library(MASS)
library(NHANES)
library(plotly)
library(ggthemes)
library(metR)
library(shiny)
#library(xlsx)
cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
suppressWarnings(library(dplyr))
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

#Research Question 1:
```{r}
world = read.csv("data/dep_share.csv") %>% filter(Year == 2017)
gender = read.csv("data/depression-by-gender.csv")
continents = read.csv("data/continent.csv")
continents$location = continents$Country
```

```{r, echo=FALSE}
world_coordinates <- map_data("world")
world_coordinates$Entity = world_coordinates$region
world_depression = left_join(world, world_coordinates, by = "Entity")
world_depression$percentage_dep = world_depression$Prevalence...Depressive.disorders...Sex..Both...Age..Age.standardized..Percent.
ggplot(world_depression, aes(x = long, y = lat, group = group, fill = percentage_dep)) +
  geom_polygon(color = "gray", size = 0.1) +
  theme_map() +
  scale_fill_gradient(low = "yellow", high = "red") +
  ggtitle("Percentage of Depression by Country
          ") +
  labs(subtitle = "Red represetns high depression and Yellow is low depression",
       caption = "Data downloaded from https://ourworldindata.org/mental-health#citation")
```

```{r}
gender_continent = left_join(gender,continents,by= "location")
gender_continent = na.omit(gender_continent)
gender_continent$val = gender_continent$val * 100

v= data.frame(Gender = recode(gender_continent$sex,"Male" = 1,"Female"=0),Depression= gender_continent$val,Continent = gender_continent$Continent)

ggplot(v,aes(x =Gender, y =Depression))+geom_jitter()+geom_smooth(method ="lm",col="orange") +facet_wrap(~Continent,nrow=2) +ylab("Depression Percentage") +xlab("Gender") + ggtitle("Continent Analysis of Depression (1990-2017) by Gender") +
  labs(subtitle = "0= Female & 1=Male",
       caption = "Data downloaded from https://ourworldindata.org/mental-health#citation")
```

```{r}
trend = gender_continent %>% group_by(Continent,year,sex) %>% summarise(avgdep = mean(val))

ggplot(trend, aes(x=year, y = avgdep,col=sex)) + geom_line() + facet_wrap(~Continent,nrow=2) +
  xlab("Years") + ylab("Average Depression") + ggtitle("Trend for depression by continents and gender")  +
  labs(caption = "Data downloaded from https://ourworldindata.org/mental-health#citation")
```

#Research Question 2:
Load mental health depression data
```{r}
depByAge <- read.csv("data/depression-by-age.csv", header = TRUE)
```

Data cleanup
```{r}
depByAge.df <- depByAge %>% 
  rename("year" = "year",
         "country" = "location",
         "depression" = "val") %>% 
  filter(age %in% c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years")) %>% 
  dplyr::select(year, country,age, depression) %>% 
  mutate(depression = depression *100)
```

Load continent <> country mapping
```{r}
continents <- read.csv("data/continent.csv", header = TRUE) %>% 
  rename("country" = "Country", "continent" = "Continent")
```

Load country wise GDP data
```{r}
gdpPerCap <- read.csv("data/GM-GDP per capita - Dataset - v28 - data-for-countries-etc-by-year.csv") %>% 
  rename("country" = "name", "year" = "time") %>% 
  dplyr::select(country, year, Income.per.person)
  
```

Load population data
```{r}
population = read.csv("data/GM-Population - Dataset - v6 - data-for-countries-etc-by-year.csv")
populationFiltered = filter(population,time>=1990 & time<=2017) %>% 
  rename("country" = "name", "year" = "time")
```

Adding continents to depression data
```{r}
depByAge.cont <- depByAge.df %>% 
  left_join(continents, by = "country") %>% 
  dplyr::select(year, continent, country, age,depression) %>% 
  filter(!is.na(continent))
```

```{r}
depwGDP <- depByAge.cont %>% left_join(gdpPerCap, by = "country")
```

Plot for depression vs age group in USA over time
```{r}
depwGDP %>% 
  group_by(continent, age) %>% 
  summarise(avg_depPct = median(depression)) %>% 
ggplot(aes(x = age, y = avg_depPct))+
  geom_point()+
  facet_wrap(~ continent)+
  scale_color_manual(values = cbPalette)+
  theme_bw()+
  ggtitle("Depression % among age groups across continents")+
  labs(subtitle = "",
       caption = "Data from https://ourworldindata.org")+
  ylab("Median Depression %")+
  xlab("Age bracket")+
    scale_x_discrete(guide = guide_axis(angle = 60)) 

```

```{r}

age.medians <- depByAge.cont %>% 
  mutate(age =recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6)) %>% 
  filter(!is.na(age)) %>% 
  group_by(age) %>% 
  dplyr::summarise(medianDep = median(depression))
ggplot(depByAge.cont, aes(x = recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6), y = depression))+
  geom_jitter()+
  geom_point(data=age.medians,aes(x = age, y = medianDep), color = "orange")+
 scale_x_continuous(name = "Age group", breaks = c(1,2,3,4,5,6), labels = c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years"),
                    guide = guide_axis(angle = 60))+
  ylab("Depression %")+
  ggtitle("Depression % in age groups")+
  labs(subtitle = "Orange dots represent median depression %",
       caption = "Data from https://ourworldindata.org")+
  theme_bw()
  
```


```{r}

age.medians <- depByAge.cont %>% 
  mutate(age =recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6)) %>% 
  filter(!is.na(age)) %>% 
  group_by(age) %>% 
  dplyr::summarise(medianDep = median(depression))
ggplot(depByAge.cont, aes(x = recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6), y = depression))+
  geom_jitter()+
  geom_point(data=age.medians,aes(x = age, y = medianDep), color = "orange")+
 scale_x_continuous(name = "Age group", breaks = c(1,2,3,4,5,6), labels = c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years"),
                    guide = guide_axis(angle = 60))+
  ylab("Depression %")+
  ggtitle("Depression % in age groups")+
  labs(subtitle = "Orange dots represent median depression %",
       caption = "Data from https://ourworldindata.org")+
  theme_bw()
  
```

Depression trend in North America
```{r}
depByAge.usa <- depByAge.cont %>% 
  filter(continent == "North America" | continent == "Oceania") %>%
  filter(country %in% c("Greenland", "Canada", "Australia", "New Zealand")) %>% 
  mutate(country = factor(country, levels = c("Greenland", "Canada", "Australia", "New Zealand"))) %>% 
  arrange(continent)
```

```{r}
depByAge.usa.medians <- depByAge.usa %>% 
  mutate(age =recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6)) %>% 
  filter(!is.na(age)) %>% 
  group_by(country,age) %>% 
  dplyr::summarise(medianDep = median(depression))
depByAge.usa %>% 
ggplot(aes(x = recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6), y = depression))+
  geom_jitter()+
  geom_point(data=depByAge.usa.medians,aes(x = age, y = medianDep), color = "orange")+
  facet_wrap(~country)+
 scale_x_continuous(name = "Age group", breaks = c(1,2,3,4,5,6), labels = c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years"),
                    guide = guide_axis(angle = 60))+
  ylab("Depression %")+
  ggtitle("Depression trend exceptions in North America and Oceania")+
  labs(subtitle = "Orange dots represent median depression %",
       caption = "Data from https://ourworldindata.org")+
  theme_bw()
  
```

Depression trend in Oceania
```{r}
depByAge.oceania <- depByAge.cont %>% 
filter(continent == "Oceania", country %in% c("Australia", "New Zealand")) 
```

```{r}
depByAge.oceania.medians <- depByAge.oceania%>% 
  mutate(age =recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6)) %>% 
  filter(!is.na(age)) %>% 
  group_by(country,age) %>% 
  dplyr::summarise(medianDep = median(depression))
depByAge.oceania %>% 
ggplot(aes(x = recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6), y = depression))+
  geom_jitter()+
  geom_point(data=depByAge.oceania.medians,aes(x = age, y = medianDep), color = "orange")+
  facet_wrap(~country)+
 scale_x_continuous(name = "Age group", breaks = c(1,2,3,4,5,6), labels = c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years"),
                    guide = guide_axis(angle = 60))+
  ylab("Depression %")+
  ggtitle("Depression % in age groups across Australia and New Zealand")+
  labs(subtitle = "Orange dots represent median depression %",
       caption = "Data from https://ourworldindata.org")+
  theme_bw()
  
```
```{r}
depByAge.recoded <- depByAge.cont%>% 
  mutate(age =recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6))
age.lm <- lm( depByAge.recoded$depression ~ depByAge.recoded$age )
display(age.lm)
AIC(age.lm)

```

Fit a model for age and sex
```{r}
depbyAgeSex <- read.csv("data/depression-by-age-gender.csv", header = TRUE)
depbyAgeSex.df <- depbyAgeSex %>% 
  rename("country" = "location",
         "depression" = "val"
         ) %>% 
  filter(age %in% c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years")) %>% 
  dplyr::select(year, country,age,sex, depression) %>% 
  mutate(depression = depression *100)
```

```{r}
depbyAgeSex.df <- depbyAgeSex.df %>% 
  mutate(age =recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6),
         sex = recode(sex, "Male" = 1, "Female"=2))

```

```{r}
age.lm <- lm(depbyAgeSex.df$depression ~ depbyAgeSex.df$age + depbyAgeSex.df$sex )
age.lm2 <- lm(depbyAgeSex.df$depression ~ depbyAgeSex.df$age + depbyAgeSex.df$sex + depbyAgeSex.df$age:depbyAgeSex.df$sex )
AIC(age.lm)
AIC(age.lm2)
display(age.lm)
display(age.lm2)
```


```{r}
model.df <- data.frame(age = depbyAgeSex.df$age,
                       sex = depbyAgeSex.df$sex,
                       .resid = residuals(age.lm2),
                       .fitted = fitted.values(age.lm2))
```

```{r}
model.grid <- expand.grid(age = c(1,2,3,4,5,6), sex = c(1,2))
dep.predict <- predict(age.lm2, newdata = model.grid)
depmodel.df <- data.frame(model.grid, depression = as.vector(dep.predict))
```

```{r}
ggplot(depmodel.df, aes(x = age, y = depression))+
  geom_point()+
  facet_wrap(~sex) +
  xlab("Age")+
  ylab("Sex")+
  ggtitle("Figure 4: Height vs Age, filled by Average BMI(log scaled)")+
  labs(subtitle = "Lines represent contours",
       fill = "Average BMI(log scaled)",
       caption="Data created from values based on NHANES dataset")+
  theme_bw()
```

```{r}
ggplot(model.df, aes(x = age, y = .resid))+
  geom_jitter()+
  geom_smooth(method = "loess", se = FALSE)
```

```{r}
ggplot(model.df, aes(x = sex, y = .resid))+
  geom_jitter()+
  geom_smooth(method = "loess", se = FALSE)

```

# Research Question 3:
```{r}
data = read.csv("data/depression-disorders.csv")
```

```{r}
continent = read.csv("data/continent.csv")
```

```{r}
filteredData = data%>%
  filter(cause=="Depressive disorders")%>%
  filter(year==2017)%>%
  select(location,val)
filteredData$location <- gsub("United States of America", "United States", filteredData$location)
```

```{r}
gdp_data = read.csv("data/API_NY.GDP.PCAP.CD_DS2_en_csv_v2_4701206.csv",skip = 4)
gdp_data = gdp_data%>%
  select(Country.Name,X2017)
```
```{r}
gdp = read.csv("data/GM-GDP per capita - Dataset - v28 - data-for-countries-etc-by-year.csv")
```

```{r}
filtered_gdp = gdp%>%
  filter(time==2017)%>%
  select("name","time","GDP.total")
```

```{r}
merged_df = gdp_data%>%
  left_join(filtered_gdp,by=c("Country.Name"="name"))%>%
  na.omit()
```

```{r}
merged_df = merged_df %>%
  left_join(filteredData,by=c("Country.Name"="location"))%>%
  na.omit()
```

```{r}
merged_df = merged_df%>%
  left_join(continent,by=c("Country.Name"="Country"))%>%
  na.omit()
```


```{r}
data%>%
  filter(cause=="Depressive disorders")%>%
  select(location,val,year)%>%
  left_join(continent,by=c("location"="Country"))%>%
  group_by(Continent,year)%>%
  summarise(Depression=mean(val))%>%
  drop_na()%>%
  ggplot(aes(x=year,y=Depression,color=Continent))+
  geom_line()+
  scale_color_manual(values=cb_palette)+
  xlab("Year")+
  ylab("Depression(%)")+
  labs(title = "Depression Rate by continents over the years 1990-2017",
              caption = "Data source: https://data.world/vizzup/mental-health-depression-disorder-data")
```

```{r}
merged_df$CountryBracket = case_when(merged_df$X2017<1006~"Low Income",
                                     merged_df$X2017>=1006 & merged_df$X2017<3955 ~"Lower-middle Income",
                                     merged_df$X2017>=3955 & merged_df$X2017<12235 ~"Upper-middle Income",
                                     merged_df$X2017>=12235 ~"High Income")
```

```{r}
merged_df$CountryBracket =  factor(merged_df$CountryBracket, levels=c("Low Income","Lower-middle Income","Upper-middle Income","High Income"))
merged_df$mul100 = merged_df$val*100
ggplot(merged_df,aes(x=GDP.total,y=mul100))+
  geom_point()+
  geom_smooth(method='lm',se=FALSE)+
  scale_x_log10()+
  facet_wrap(~CountryBracket)+
  # scale_color_manual(values=cb_palette)+
  xlab("GDP of the country")+
  ylab("Depression(%)")+
  labs(title = "Depression Rate by total GDP of the countries",
              subtitle = "Faceted by GDP per Capita income group",
              caption = "Data source: https://data.world/vizzup/mental-health-depression-disorder-data")
```

```{r}
ggplot(merged_df,aes(x=X2017,y=mul100))+
geom_point()+
geom_smooth(color="orange")+
scale_x_log10()
```

```{r}
ggplot(merged_df,aes(x=X2017,y=mul100))+
geom_point(size=2)+
geom_smooth(method="gam",se=FALSE)+
scale_x_log10()+
xlab("GDP Per Capita")+
ylab("Depression(%)")+
labs(title = "Figure 3.1 Depression Rate by GDP Per Capita",
            subtitle = "Blue line represents the fitted GAM model",
            caption = "Data from: https://ourworldindata.org\n
                      https://data.worldbank.org")


ggplot(merged_df,aes(x=X2017,y=mul100))+
geom_point(size=2)+
geom_smooth(method="lm",se=FALSE)+
scale_x_log10()+
xlab("GDP Per Capita")+
ylab("Depression(%)")+
scale_color_manual(values=cb_palette)+
facet_wrap(~Continent)+
labs(title = "Depression Rate by GDP Per Capita Faceted by Continents",
            subtitle = "Blue line represents the overall trend",
            caption = "Data from: https://ourworldindata.org\n
                      https://data.worldbank.org")
```

```{r}
ggplot(merged_df,aes(x=X2017,y=mul100))+
geom_point(aes(color=CountryBracket),size=2)+
geom_smooth(method="lm",se=FALSE)+
scale_x_log10()+
xlab("GDP Per Capita")+
ylab("Depression(%)")+
scale_color_manual(values=cb_palette)+
facet_wrap(~Continent)+
labs(title = "Figure 3.2 Depression Rate by GDP Per Capita Faceted by Continents",
            subtitle = "Blue line represents the overall trend",
            caption = "Data from : https://ourworldindata.org\n
            https://data.worldbank.org",
     color='Income Group')
```


```{r}
merged_df$logX2017 = log10(merged_df$X2017)
depression.gam = gam(mul100~s(logX2017),data=merged_df,method="REML")
depression.gam.df = augment(depression.gam)
```

```{r}
ggplot(depression.gam.df, aes(x=logX2017,y=.resid)) + geom_jitter() + geom_smooth(method = "loess") +
    xlab("Alcohol Disorder population (in percent)") +
  ggtitle('Residual plot with alcohol disorder population') +
  labs(subtitle = "GAM Model",
    caption = "Data domain : Mental health")
```

```{r}
ggplot(depression.gam.df,aes(x=.fitted,y=abs(.resid))) + geom_point() + geom_smooth(method="lm")
```

```{r}
summary(depression.gam)
```

```{r}
newdata1 = expand.grid(X2017=seq(250,15000,100))
newdata2 = expand.grid(X2017=seq(15500,150000,1000))
newdata = combine(newdata1$X2017,newdata2$X2017)
newdata = data.frame(X2017=newdata)
newdata$logX2017 = log10(newdata$X2017)
```

```{r}
depression.pred = predict(depression.gam,newdata=newdata)
depression.pred = data.frame(newdata,predicted_depression=as.vector(depression.pred))
```


```{r}
ggplot(depression.pred,aes(x=X2017,y=predicted_depression))+
geom_point()+
# geom_smooth(color="orange")+
scale_x_log10()
```

## Research Question 4:
```{r}
continents <- read.csv("data/continent.csv", header = TRUE) %>% 
  rename("country" = "Country", "continent" = "Continent")

data <- read.csv("data/depression-disorders.csv",sep=',') %>% drop_na() %>% rename("country" = "location") %>% inner_join(continents,by="country") %>% dplyr::select(-c("upper","lower","metric","measure","sex","age")) %>% reshape(idvar = c("country","year","continent"), timevar = "cause", direction = "wide") %>% clean_names() %>% select(-c("country","continent"))

```

```{r}
ggpairs(data[c('val_alcohol_use_disorders','val_drug_use_disorders','val_depressive_disorders')])
```

Distribution of each column for year 2000
```{r}
#alcohol_use_disorders, drug_use_disorders
data_2000 <- read.csv("data/depression-disorders.csv",sep=',') %>% drop_na() %>% rename("country" = "location") %>% filter(year == '2000') %>% inner_join(continents,by="country") %>% dplyr::select(-c("upper","lower","metric","measure","sex","age")) %>% reshape(idvar = c("country","year","continent"), timevar = "cause", direction = "wide") %>% clean_names() %>% select(c("continent","val_alcohol_use_disorders","val_drug_use_disorders","val_depressive_disorders")) %>% mutate_at(vars(c("val_alcohol_use_disorders","val_drug_use_disorders","val_depressive_disorders")), .funs = funs(. * 100))

ggplot(data_2000,aes(x=val_alcohol_use_disorders,y=val_depressive_disorders,color = continent)) +geom_point() +  scale_color_manual(values=cbPalette) + geom_smooth(method = "loess",se=FALSE, color="black") +
  xlab("Alcohol Disorder population (in percent)") +
  ylab("Depressed population (in percent)")+
  ggtitle('Depression as a function of alcohol disorder in year 2000') +
  labs(caption = "Data domain : Mental health")

ggplot(data_2000,aes(x=val_drug_use_disorders,y=val_depressive_disorders,color = continent)) +geom_point() +  scale_color_manual(values=cbPalette) + geom_smooth(method = "loess",se=FALSE, color="black") +
  xlab("Drug Disorder population (in percent)") +
  ylab("Depressed population (in percent)")+
  ggtitle('Depression as a function of drug disorder in 2000') +
  labs(caption = "Data domain : Mental health")


ggplot(data_2000,aes(x=val_alcohol_use_disorders,y=val_drug_use_disorders,color = continent)) + geom_jitter() +
  scale_color_manual(values=cbPalette) +
  xlab("Alcohol Disorder population (in percent)") +
  ylab("Drug population (in percent)")+
  ggtitle('Alcohol vs Drug') +
  labs(caption = "Data domain : Mental health")

```

Main data
```{r}
m_data <- read.csv("data/depression-disorders.csv",sep=',') %>% drop_na() %>% rename("country" = "location") %>% inner_join(continents,by="country") %>% dplyr::select(-c("upper","lower","metric","measure","sex","age")) %>% reshape(idvar = c("country","year","continent"), timevar = "cause", direction = "wide") %>% clean_names() %>% select(c("continent","year","val_alcohol_use_disorders","val_drug_use_disorders","val_depressive_disorders")) %>% mutate_at(vars(c("val_alcohol_use_disorders","val_drug_use_disorders","val_depressive_disorders")), .funs = funs(. * 100))
```

Linear model
```{r}
gap.lm2 = lm(val_depressive_disorders ~ val_alcohol_use_disorders, data = m_data)
display(gap.lm2)

gap.lm.df <- data.frame(m_data,
                        .fitted = fitted.values(gap.lm2),
                        .resid = residuals(gap.lm2))

data.grid2 <- expand.grid(val_alcohol_use_disorders = 0.25:3.5,val_drug_use_disorders = 0.05:1.5)
data.pred = predict(gap.lm2,newdata = data.grid2)
data.grid2$depression_pred = as.vector(data.pred)
data.plot.df <- data.frame(data.grid2)

ggplot(data.plot.df, aes(x = val_alcohol_use_disorders, y = depression_pred)) + geom_point() 
ggplot(data.plot.df, aes(x = val_drug_use_disorders, y = depression_pred)) + geom_point()

```


```{r}
gap.lm2 = lm(val_depressive_disorders ~ val_drug_use_disorders, data = m_data)
display(gap.lm2)

gap.lm.df <- data.frame(m_data,
                        .fitted = fitted.values(gap.lm2),
                        .resid = residuals(gap.lm2))

data.grid2 <- expand.grid(val_alcohol_use_disorders = 0.25:3.5,val_drug_use_disorders = 0.05:1.5)
data.pred = predict(gap.lm2,newdata = data.grid2)
data.grid2$depression_pred = as.vector(data.pred)
data.plot.df <- data.frame(data.grid2)

ggplot(data.plot.df, aes(x = val_alcohol_use_disorders, y = depression_pred)) + geom_point() 
ggplot(data.plot.df, aes(x = val_drug_use_disorders, y = depression_pred)) + geom_point()

```


```{r}
gap.lm2 = lm(val_depressive_disorders ~ val_alcohol_use_disorders+val_drug_use_disorders, data = m_data)
display(gap.lm2)

gap.lm.df <- data.frame(m_data,
                        .fitted = fitted.values(gap.lm2),
                        .resid = residuals(gap.lm2))

data.grid2 <- expand.grid(val_alcohol_use_disorders = 0.25:3.5,val_drug_use_disorders = 0.05:1.5)
data.pred = predict(gap.lm2,newdata = data.grid2)
data.grid2$depression_pred = as.vector(data.pred)
data.plot.df <- data.frame(data.grid2)

ggplot(data.plot.df, aes(x = val_alcohol_use_disorders, y = depression_pred)) + geom_point() 
ggplot(data.plot.df, aes(x = val_drug_use_disorders, y = depression_pred)) + geom_point()

```

Linear model
```{r}
gap.lm2 = lm(val_depressive_disorders ~ val_alcohol_use_disorders + val_drug_use_disorders + val_alcohol_use_disorders:val_drug_use_disorders, data = m_data)
display(gap.lm2)

gap.lm.df <- data.frame(m_data,
                        .fitted = fitted.values(gap.lm2),
                        .resid = residuals(gap.lm2))

data.grid2 <- expand.grid(val_alcohol_use_disorders = 0.25:3.5,val_drug_use_disorders = 0.05:1.5)
data.pred = predict(gap.lm2,newdata = data.grid2)
data.grid2$depression_pred = as.vector(data.pred)
data.plot.df <- data.frame(data.grid2)

ggplot(data.plot.df, aes(x = val_alcohol_use_disorders, y = depression_pred)) + geom_point() 
ggplot(data.plot.df, aes(x = val_drug_use_disorders, y = depression_pred)) + geom_point()

```


#Research Question 5
Final model:
```{r}
genagedis = read.csv("data/depression-age-gender-disorders.csv")
data = genagedis %>% dplyr::select(sex,age,cause,val,year,location)
z = data %>% tidyr::pivot_wider(values_from = "val",names_from = "cause")
z$drug = z$`Drug use disorders` *100
z$dep = z$`Depressive disorders` * 100
z$alcohol = z$`Alcohol use disorders` *100

z= na.omit(z)
z = z %>% filter(age %in% c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years"))

depbyAgeGenDis <- z %>% 
  mutate(age =recode(age, "10-14 years" = 1,   "15-19 years"=2, "25-29 years"=3, "30-34 years"=4,  "50-69 years"=5, "70+ years"=6),
         sex = recode(sex, "Male" = 1, "Female"=2))

dep.lm = lm(dep ~ age + sex+ age:sex+alcohol+age:alcohol+sex:alcohol+drug+age:drug+sex:drug ,data = depbyAgeGenDis)
display(dep.lm)
```

```{r}
colnames(data)
```


```{r}
model.df <- data.frame(age = depbyAgeGenDis$age,
                       sex = depbyAgeGenDis$sex,
                       alcohol = depbyAgeGenDis$alcohol,
                       drug = depbyAgeGenDis$drug,
                       .resid = residuals(dep.lm),
                       .fitted = fitted.values(dep.lm))
```

```{r}
model.grid <- expand.grid(age = c(1,2,3,4,5,6), sex = c(1,2), alcohol = seq(0.1, 17, 0.5), drug = seq(0.1, 6, 0.5))
dep.predict <- predict(dep.lm, newdata = model.grid)
depmodel.df <- data.frame(model.grid, depression = as.vector(dep.predict))
```

```{r}
sexes <- c(`1` = "Male",
           `2` = "Female")
ggplot(depmodel.df, aes(x = alcohol, y = depression, group = age, color = as.factor(age)))+
  geom_smooth(se = FALSE)+
  facet_wrap(~sex, labeller = as_labeller(sexes))+
  xlab("Alcohol use %")+
  ylab("Depression %")+
  ggtitle("Depression  % vs Alcohol use % faceted by sex")+
  labs(subtitle = "Colors of lines represent different age groups",
       caption="Data from https://ourworldindata.org",
       color = "Age Group"
      )+
  scale_color_manual(labels = c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years"),
                     values = cbPalette)+
  theme_bw()
```

```{r}
sexes <- c(`1` = "Male",
           `2` = "Female")
ggplot(depmodel.df, aes(x = drug, y = depression, group = age, color = as.factor(age)))+
  geom_smooth(se = FALSE)+
  facet_wrap(~sex, labeller = as_labeller(sexes))+
  xlab("Drug use %")+
  ylab("Depression %")+
  ggtitle("Depression  % vs Drug use % faceted by sex")+
  labs(subtitle = "Colors of lines represent different age groups",
       caption="Data from https://ourworldindata.org",
       color = "Age Group"
      )+
  scale_color_manual(labels = c("10-14 years",  "15-19 years","25-29 years","30-34 years","50-69 years","70+ years"),
                     values = cbPalette)+
  theme_bw()
```

